\chapter*{Acknowledgements}
We’ve reached the end, not just of this thesis but of the entire journey.
Now that I've arrived, I feel I have some thank-yous to give.

First and foremost, I must thank my supervisor, who, despite numerous commitments, guided me to the end, allowing me to complete this work. I challenged myself and pushed beyond what I thought I was capable of, and I owe this achievement to him.

There is no knowledge I’ve gained in these three years that I couldn’t have learned on my own from some online article or textbook. For this reason, a huge thank you goes to my classmates. You are the true added value of this university. Genuine companions, always willing to help others out of pure altruism.
There is no shame or exaggeration in saying that without you, I wouldn’t have gotten to where I am now. The exchange and interaction we’ve had has been the unique and indispensable resource that enriched this journey. Together, we encouraged each other and forged our way along a path that was often uphill but, in the end, always rewarded the effort.

A special thank you goes to the friends who have stood by me during the writing of this thesis and throughout this past year. You have given me so much, and I hope I can give back as much as I have received.

Thanks to my lifelong friends, who have stayed close to me even when I’ve been difficult or had my head elsewhere. Your support, lightness, and spontaneity remind me that we are not made of just flesh and rationality.

I have two more thank yous to give: The first goes to my girlfriend, Sofia. \\I will never find the right words to thank you for everything. Thank you for every moment we’ve shared; seeing things from your perspective continues to improve and surprise me.
Lastly, I must thank my parents, who push me every day to give my best because they believe I can do far more and better than I think I can (and, in the end, maybe they are right).

Thanks again to everyone.

\addcontentsline{toc}{chapter}{Acknowledgements}
